Approximation as Differentiation notes

### 2025-03-10 notes/sec/grothendieck.tex
Correction to isomorphism property.
Indexing function for lambda inv.
Indexing component of lambda is an iso by Cartesian closure of Set.

\begin{enumerate}
\item $[X, Y]$ to be the object in $\Fam(C)$ with $\idx{[X,Y]} = \Fam(C)(X,Y)$ and $\faml{[X, Y]}(f) =
\prod_{\idx{X}}\reindex{\faml{Y}}{\idx{f}}$ for every $f: X \to Y$.
\item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) sending any $f: X \times Y \to Z$ to
$\lambda_{X,Y,Z}(f): X \to [Y, Z]$ with:
\begin{itemize}
\item $\idx{\lambda}_{X,Y,Z}(f)(x) = g$ where $\idx{g} = \idx{f} \comp (x, -): \idx{Y} \to \idx{Z}$ and
$\faml{g} = \reindex{\faml{f}}{(x,-)} \comp \inj_{\faml{Y}}: \faml{Y} \to \reindex{\faml{Z}}{\idx{g}}$ where $\inj_{\faml{Y}}:
\faml{Y} \to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
\item $\faml{\lambda}_{X,Y,Z}(f)(x) = \lambda_{\prod_{\idx{Y}}}(\reindex{\faml{f}}{(x,-)} \comp \inj_{\faml{X}_x}:
\const_{\idx{Y}}(\faml{X}_x) \to \reindex{\faml{Z}}{\idx{f}})$ where $\inj_{\faml{X}_x}: \const_{\idx{Y}}(\faml{X}_x)
\to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
\end{itemize}
\item The family of morphisms $\eval_{X,Y}: [X, Y] \times X \to Y$ in $\Fam(C)$ with
\begin{itemize}
\item $\idx{\eval}_{X,Y}(f, x) = \idx{f}(x)$
\item $\faml{\eval}_{X,Y}(f, x) = \coprodM{{\eval_{\prod}}_x}{\faml{f}_x}:
\prod_{\idx{X}}\reindex{\faml{Y}}{\idx{f}} \biprod \faml{X}_x \to \faml{Y}_{\idx{f}(x)}$
\end{itemize}
using $\coprodM{-}{-}$ to denote coproduct of morphisms.
\end{enumerate}
Then for any $f: X \times Y \to Z$ in $\Fam(C)$:
\begin{enumerate}
\item Suppose $(x, y) \in \idx{X} \times \idx{Y}$. The diagram on the left below commutes by the universal property
of the set-indexed product; the commutativity of the diagram on the right is trivial.

@@

\begin{center}
\begin{tikzcd}[column sep=5.5em]
   \prod_{\idx{Y}}\reindex{\faml{Z}}{\idx{f} \comp (x,-)} \biprod \faml{Y}_y \arrow[dr,
   "\coprodM{\eval_{\prod_{y}}}{\faml{f}_{(x,y)}\,\comp\,\inj_{\faml{Y}_y}}"] \\
   \faml{X}_x \biprod \faml{Y}_y \arrow[u, "(\lambda_{\prod_{\idx{Y}}}(\reindex{\faml{f}}{{(x,-)}}\,\comp\,
\inj_{\faml{X}_x}){,}\,\id_{\faml{Y}_y})"] \arrow[r, "\coprodM{\faml{f}_{x,y}\,\comp\,\inj_{\faml{X}_x}}{\faml{f}_{x,y}\,\comp\,\inj_{\faml{Y}_y}}"'] & \faml{Z}_{\idx{f}(x,y)}
\end{tikzcd}
\hspace{1cm}
\begin{tikzcd}
   {{[Y, Z]}} \tensor Y \arrow[dr, "\eval_{Y,Z}"] \\
   X \tensor Y \arrow[u, "\lambda_{X,Y,Z}(f) \tensor \id_Y"] \arrow[r, "f"'] & Z
\end{tikzcd}
\end{center}

\item \todo{(exhibit $\lambda^{-1}$ and then show) $\lambda^{-1}_{X,Y,Z}(\lambda_{X,Y,Z}(f)) = f$}
\end{enumerate}
\end{proof}

>>>>

\begin{enumerate}
\item $[X, Y]$ to be the object in $\Fam(C)$ with $\idx{[X,Y]} = \Fam(C)(X,Y)$ and $\faml{[X, Y]}(f) =
\prod_{\idx{X}}\reindex{\faml{Y}}{\idx{f}}$ for every $f: X \to Y$.
\item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) sending any $f: X \times Y \to Z$ to
$\lambda_{X,Y,Z}(f): X \to [Y, Z]$ with:
\begin{itemize}
\item $\idx{\lambda}_{X,Y,Z}(f)(x) = g$ where $\idx{g} = \idx{f} \comp (x, -): \idx{Y} \to \idx{Z}$ and
$\faml{g} = \reindex{\faml{f}}{(x,-)} \comp \inj_{\faml{Y}}: \faml{Y} \to \reindex{\faml{Z}}{\idx{g}}$ where $\inj_{\faml{Y}}:
\faml{Y} \to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
\item $\faml{\lambda}_{X,Y,Z}(f)(x) = \lambda_{\prod_{\idx{Y}}}(\reindex{\faml{f}}{(x,-)} \comp \inj_{\faml{X}_x}:
\const_{\idx{Y}}(\faml{X}_x) \to \reindex{\faml{Z}}{\idx{f}})$ where $\inj_{\faml{X}_x}: \const_{\idx{Y}}(\faml{X}_x)
\to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
\end{itemize}
\item The family of morphisms $\eval_{X,Y}: [X, Y] \times X \to Y$ in $\Fam(C)$ with
\begin{itemize}
\item $\idx{\eval}_{X,Y}(f, x) = \idx{f}(x)$
\item $\faml{\eval}_{X,Y}(f, x) = \coprodM{{\eval_{\prod}}_x}{\faml{f}_x}:
\prod_{\idx{X}}\reindex{\faml{Y}}{\idx{f}} \biprod \faml{X}_x \to \faml{Y}_{\idx{f}(x)}$
\end{itemize}
using $\coprodM{-}{-}$ to denote coproduct of morphisms.
\end{enumerate}

Then:

\begin{enumerate}
\item $\idx{\lambda}_{X,Y,Z}$ is an isomorphism by the ismorphism of $\idx{X} \times \idx{Y} \to \idx{Z}$ and
$\idx{X} \to \idx{Y} \to \idx{Z}$. For $\faml{\lambda}_{X,Y,Z}$, define the family of functions
$\lambda^{-1}_{X,Y,Z}$ sending any $f: X \to [Y,Z]$ to $\lambda^{-1}_{X,Y,Z}(f): X \times Y \to Z$ with:
\begin{itemize}
\item $\idx{\lambda^{-1}}_{X,Y,Z}(f)(x,y) = \idx{g}(y)$ where $\idx{f}(x) = g$
\item $\faml{\lambda^{-1}}_{X,Y,Z}(f)(x,y) =$
\end{itemize}

\todo{Show $\lambda^{-1}_{X,Y,Z}(\lambda_{X,Y,Z}(f)) = f$ and $\lambda_{X,Y,Z}(\lambda^{-1}_{X,Y,Z}(f)) =
f$}.

\item Suppose $f: X \times Y \to Z$ in $\Fam(C)$ and $(x, y) \in \idx{X} \times \idx{Y}$. The diagram on the
left below commutes by the universal property of the set-indexed product; the commutativity of the diagram on
the right is trivial.

@@

\begin{center}
\begin{tikzcd}[column sep=5.5em]
   \prod_{\idx{Y}}\reindex{\faml{Z}}{\idx{f} \comp (x,-)} \biprod \faml{Y}_y \arrow[dr,
   "\coprodM{\eval_{\prod_{y}}}{\faml{f}_{(x,y)}\,\comp\,\inj_{\faml{Y}_y}}"] \\
   \faml{X}_x \biprod \faml{Y}_y \arrow[u, "(\lambda_{\prod_{\idx{Y}}}(\reindex{\faml{f}}{{(x,-)}}\,\comp\,
\inj_{\faml{X}_x}){,}\,\id_{\faml{Y}_y})"] \arrow[r,
"\coprodM{\faml{f}_{x,y}\,\comp\,\inj_{\faml{X}_x}}{\faml{f}_{x,y}\,\comp\,\inj_{\faml{Y}_y}}"'] &
\faml{Z}_{\idx{f}(x,y)}
\end{tikzcd}
\hspace{1cm}
\begin{tikzcd}
   {{[Y, Z]}} \tensor Y \arrow[dr, "\eval_{Y,Z}"] \\
   X \tensor Y \arrow[u, "\lambda_{X,Y,Z}(f) \tensor \id_Y"] \arrow[r, "f"'] & Z
\end{tikzcd}
\end{center}
\end{enumerate}
\end{proof}
